---
title: "Day 2 - Lines"
format: html
editor: visual
---

# Load packages

```{r}
#| message: false
library(tidyverse) # Tidyverse pkgs
library(sf) # Work with spatial data
library(showtext) # Custom fonts
library(cowplot) # For combining plot and insets/text
library(ggtext) # Markdown text
library(htmltools) # Custom font styles
library(osmdata) # OSM data
library(geodata) # To get elevation data
library(rnaturalearthhires) # High res border shapefile
library(stars) # Working with raster data
library(ggridges) # To make ridge plot

showtext::showtext_auto()
showtext_opts(dpi = 600) # Makes sure the text is same res as plot
## https://fonts.adobe.com/fonts/xenara
sysfonts::font.add(family = "Xenara",
                   regular = "Xenara_regular.otf",
                   bold = "Xenara_bold.otf")
```

```{r}
# Load detailed state boundary data
states <- rnaturalearthhires::states10
```

```{r}
# Load detailed elevation data from around Canberra
elev <- geodata::elevation_3s(lon = 149.128998, lat = -35.282001,
                              path = here::here("Day2"))
```

Isolate ACT as an sf object

```{r}
ACT <- states %>% 
  sf::st_as_sf() %>% 
  filter(name == "Australian Capital Territory") 
```

Extract bounding area of the polygon

```{r}
(ACT_bbox <- sf::st_bbox(ACT))
```

Create lines across ACT at evenly spaced latitudes

```{r}
ACT_lines <- seq(ACT_bbox$ymin, ACT_bbox$ymax, length.out = 60)
```

Turn these lines into sf objects

```{r}
ACT_lines_sf <- purrr::map(ACT_lines,
           .f = ~{
             pts <- matrix(c(ACT_bbox$xmin, ACT_bbox$xmax, .x, .x), ncol = 2)
             line <- sf::st_linestring(pts, dim = "XY")
           }) %>% 
  sf::st_as_sfc(crs = "EPSG:4326")

ggplot() +
  geom_sf(data = ACT_lines_sf) +
  geom_sf(data = ACT) +
  theme_void()
```

Convert these lines into points at even spaces (where we can extract elevation)

```{r}
ACT_pts_sf <- ACT_lines_sf %>% 
  sf::st_segmentize(dfMaxLength = 500) %>% 
  sf::st_as_sf() %>% 
  sf::st_cast(to = "POINT") %>% 
  sf::st_filter(ACT)

ggplot() +
  geom_sf(data = ACT) +
  geom_sf(data = ACT_pts_sf, size = 0.1) +
  theme_void()
```

Convert raster to stars (works better with sf objects)

```{r}
elev_stars <- stars::st_as_stars(elev)
```

Extract elevation information at each point

```{r}
ACT_pts_elev <- ACT_pts_sf %>% 
  dplyr::bind_cols(stars::st_extract(elev_stars, at = .) %>% sf::st_drop_geometry()) %>% 
  dplyr::bind_cols(sf::st_coordinates(.)) %>% 
  mutate(Y = round(Y, 2),
         elev = srtm_66_20.tif - min(srtm_66_20.tif))
```

Location of Canberra to add as a point.

```{r}
canberra <- data.frame(X = 149.1245965, Y = -35.2878983)
```

Download border of Namadgi NP from OSM

```{r}
ACT_bbox <- getbb(place_name = "Australian Capital Territory, Australia")
allnature <- opq(bbox = ACT_bbox, timeout = 120) %>% 
  add_osm_feature(key = 'leisure', value = 'nature_reserve') %>% 
  osmdata_sf()

namadgi <- allnature$osm_multipolygons %>% 
  filter(name == "Namadgi National Park")
```

Subset points that are within Namadgi

```{r}
namadgi_pts <- ACT_pts_elev %>% 
  sf::st_filter(namadgi)
```

Create ridgeplot/joyplot

```{r}
forest_colour <- "#228B22"

baseplot <- ggplot() +
  geom_density_ridges(data = ACT_pts_elev, aes(x = X, y = Y, height = elev, group = Y),
                 colour = "white", fill = "black",
                 stat = "identity", scale = 8,
                 size = 0.25) +
  geom_density_ridges(data = namadgi_pts, aes(x = X, y = Y, height = elev, group = Y),
                 colour = forest_colour, fill = "black",
                 stat = "identity", scale = 8,
                 size = 0.25) +
  geom_segment(data = canberra,
               aes(x = X, xend = X, y = Y, yend = Y + 0.2),
               size = 0.25, colour = "white") +
  geom_segment(data = canberra,
               aes(x = X, xend = X + 0.15, y = Y + 0.2, yend = Y + 0.2),
               size = 0.25, colour = "white") +
  geom_richtext(data = canberra,
                aes(x = X + 0.05, y = Y + 0.23),
                label = "CANBERRA", hjust = 0, size = 5,
                colour = "white",
                fill = NA, label.colour = NA,
                family = "Xenara", fontface = "bold") +
  geom_point(data = canberra,
               aes(x = X, y = Y),
               size = 3, colour = "white") +
  theme_void() +
  theme(plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        plot.margin = margin(t = 30, b = 30,
                             l = 30, r = 30))
```

```{r}
namadgi_fn <- function(...) span(..., style = css(color = forest_colour))

ggdraw() +
  draw_plot(baseplot) +
  geom_richtext(aes(x = 0.625, y = 0.425),
                label = paste0("The Australian Capital Territory<br>is home to Australia's capital, Canberra.<br>The ACT is only 2300km^2 but alot<br>of that space is mountainous<br>", namadgi_fn("Namadgi National Park.")),
                label.colour = NA, fill = NA,
                colour = "white", family = "Xenara",
                hjust = 0, size = 3) +
  labs(title = "Australian Capital Territory",
       subtitle = "Elevation map",
       caption = "Plot: @ldbailey255 | Data: SRTM 3s elevation data") +
  theme_void() +
  theme(plot.title = element_markdown(colour = "white",
                                      family = "Xenara", size = 25,
                                      face = "bold",
                                      margin = margin(t = 20, l = 20)),
        plot.subtitle = element_markdown(colour = "white",
                                      family = "Xenara", size = 15,
                                      margin = margin(t = 5, l = 20,
                                                      b = -10)),
        plot.caption = element_markdown(colour = "white",
                                        family = "Xenara", size = 5,
                                        margin = margin(b = 5, r = 5)),
        panel.background = element_rect(colour = NA, fill = "black"),
        plot.background = element_rect(colour = NA, fill = "black"))

ggsave(here::here("./Day2/Day2_lines.png"), dpi = 600,
       height = 5, width = 8)
```
